Variational Monte Carlo on Bosonic Systems

### 2018-04-06 project_1/article/article.tex
Spell check results from importance sampling

        \subsubsection{Introducing Importance Sampling}

            Next we will introduce importance sampling, the first prediction for
            importance sampling compared to brute, naïve method is that we
            should approach an equilibrium much faster because we are now making
            moves in a smarter way. This prediction is confirmed in
            \autoref{fig:problem_c1} and \autoref{fig:problem_c2}.

            \begin{figure}
                \centering
                    \includegraphics[width=244px]
                    {../data/figures/problem_c1.pdf}
                    \caption{As the number of Monte Carlo cycles increases the
                    standard deviation of the local energies decrease. This
                    effect is stronger when the importance sampling scheme is
                    employed.}
                    \label{fig:problem_c1}
            \end{figure}

            \begin{figure}
                \centering
                    \includegraphics[width=244px]
                    {../data/figures/problem_c2.pdf}
                    \caption{The absolute difference between the expected energy
                    and the exact energy for both brute force and importance
                    sampling. }
                    \label{fig:problem_c2}
            \end{figure}

            \autoref{fig:problem_c1} shows the standard deviation of the local
            energies for increasing number of Monte Carlo cycles for both the
            regular sampling and the importance sampling methods. We see that
            the standard deviation decreases for both methods for a larger
            number of Monte Carlo cycles, but the standard deviation is always
            lower for the importance sampling method. This is in accordance with
            our expectations.

            We have also compared the expected energy with the exact energy for
            different number of Monte Carlo cycles. The result of this analysis
            is shown in \autoref{fig:problem_c2}. We see that both methods will
            evnetually result in expected energies that are equal to the exact
            energies as $\abs{E - \expval{E}} \to 0$. However, for the method of
            importance sampling the error is consistently lower.

            \begin{table}
                \caption{Importance sampled simulations of $n=10$ particles in
                $d=3$ dimensions. Average acceptance ratios and standard
                deviations for eleven alpha values between $0.3$ and $0.7$.}
                \centering
                    \begin{ruledtabular}
                        \begin{tabular}{cccc}
                            $\delta t$ & $\bar{\sigma}_b$ & $\bar{A}$
                            & $\expval{E}\left(\alpha = 1/2\right)$ \\
                            \hline
                            $2^{+3}$ & 0.09363 & 0.005 & 15.000 \\
                            $2^{+2}$ & 0.01549  & 0.041 & 15.000 \\
                            $2^{+1}$ & 0.00706 & 0.216 & 15.000 \\
                            $2^{\ 0}$ & 0.00675 & 0.510 & 15.000 \\
                            $2^{-1}$ & 0.00932 & 0.738 & 15.000 \\
                            $2^{-2}$ & 0.01578 & 0.866 & 15.000 \\
                            $2^{-3}$ & 0.02723 & 0.933 & 15.000 \\
                            $2^{-4}$ & 0.05016 & 0.966 & 15.000 \\
                            $2^{-5}$ & 0.08474 & 0.983 & 15.000 \\
                            $2^{-6}$ & 0.11555 & 0.992 & 15.000 \\
                            $2^{-7}$ & 0.17070 & 0.996 & 15.000 \\
                        \end{tabular}
                    \end{ruledtabular}
                    \label{tab:imp_time_step}
            \end{table}

            \autoref{tab:imp_time_step} shows the standard deviations and
            acceptance ratios for different time steps. The simulations in this
            tabel are from the same span $\alpha \in [0.3, 07]$ as in
            \autoref{tab:1D1N_analytic} and \autoref{tab:1D1N_numeric}. The
            optimum energies for $\alpha = \frac{1}{2}$ gives the same result
            for importance sampling as for the naïve way of sampling. Notice
            that by lowering the time step, the uncertainty starts to decrease
            and then starts to increase again. Moreover, the acceptance ratios
            increases as the time step decreases.

>>>>

        \subsubsection{Introducing Importance Sampling}
            Next we will introduce importance sampling. The first prediction for
            importance sampling compared to the brute force, naïve method is
            that we should approach an equilibrium much faster because we are
            now making moves in a smarter way. This prediction is confirmed in
            \autoref{fig:problem_c1} and \autoref{fig:problem_c2}.

            \begin{figure}
                \centering
                    \includegraphics[width=244px]
                    {../data/figures/problem_c1.pdf}
                    \caption{As the number of Monte Carlo cycles increases, the
                    standard deviation of the expectation value of the energy
                    for the brute force and importance sampling schemes
                    decreases.}
                    \label{fig:problem_c1}
            \end{figure}

            \begin{figure}
                \centering
                    \includegraphics[width=244px]
                    {../data/figures/problem_c2.pdf}
                    \caption{The absolute difference between the expected energy
                    and the exact energy for both the brute force method and
                    importance sampling.}
                    \label{fig:problem_c2}
            \end{figure}

            In \autoref{fig:problem_c1} we can see the standard deviation of the
            local energies for an increasing number of Monte Carlo cycles for
            both the brute force sampling and the importance sampling methods.
            We see that the standard deviation decreases for both methods for a
            larger number of Monte Carlo cycles, but the standard deviation is
            always lower for the importance sampling method. This is in
            accordance with our expectations.

            We have also compared the expected energy with the exact energy for
            a different number of Monte Carlo cycles. The result of this
            analysis is shown in \autoref{fig:problem_c2}. We see that both
            methods will eventually result in expected energies that are equal
            to the exact energies as $\abs{E - \expval{E}} \to 0$. However, for
            the method of importance sampling the error is consistently lower.

            \begin{table}
                \caption{Importance sampled simulations of $N=10$ particles in
                $d=3$ dimensions. Average acceptance ratios and standard
                deviations for eleven values of $\alpha \in [0.3, 0.7]$.}
                \centering
                    \begin{ruledtabular}
                        \begin{tabular}{cccc}
                            $\delta t$ & $\bar{\sigma}_b$ & $\bar{A}$
                            & $\expval{E}\left(\alpha = 1/2\right)$ \\
                            \hline
                            $2^{+3}$ & 0.09363 & 0.005 & 15.000 \\
                            $2^{+2}$ & 0.01549  & 0.041 & 15.000 \\
                            $2^{+1}$ & 0.00706 & 0.216 & 15.000 \\
                            $2^{\ 0}$ & 0.00675 & 0.510 & 15.000 \\
                            $2^{-1}$ & 0.00932 & 0.738 & 15.000 \\
                            $2^{-2}$ & 0.01578 & 0.866 & 15.000 \\
                            $2^{-3}$ & 0.02723 & 0.933 & 15.000 \\
                            $2^{-4}$ & 0.05016 & 0.966 & 15.000 \\
                            $2^{-5}$ & 0.08474 & 0.983 & 15.000 \\
                            $2^{-6}$ & 0.11555 & 0.992 & 15.000 \\
                            $2^{-7}$ & 0.17070 & 0.996 & 15.000 \\
                        \end{tabular}
                    \end{ruledtabular}
                    \label{tab:imp_time_step}
            \end{table}

            \autoref{tab:imp_time_step} shows the standard deviations and
            acceptance ratios for different time steps, $\delta t$ using the
            importance sampling method. The simulations in this table are from
            the same values of $\alpha$ as in \autoref{tab:1D1N_analytic} and
            \autoref{tab:1D1N_numeric}. The optimum energies for $\alpha = 1/2$
            gives the same result for importance sampling as for the naïve way
            of sampling. Notice that by lowering $\delta t$, the uncertainty
            starts to decrease and then starts to increase again. Moreover, the
            acceptance ratios increases as $\delta t \to 0$. This happens as the
            change in the wavefunction grows minuscule and $q(\vf{r}_{i + 1},
            \vf{r}_i) \approx 1$.
